\documentclass[tikz, a4paper]{nmd/hw}
\begin{document}
\classtemplate{Nathan's HW}{Dehn surgeon school}{\hfill}

References and links for the second computational session, which
demonstrated a wide variety of tools for studying Dehn surgery,
hyperbolic geometry, Floer homology, foliations, character varieties,
and Heegaard splittings, all of which can be used together via
Python/SageMath via the computop/sage Docker image.

\begin{itemize}[itemsep=-5pt]
\item \url{http://snappy.computop.org}
\item \url{http://www.sagemath.org}
\item \url{http://bitbucket.org/t3m/sagedocker}
\item \url{http://doi.org/10.7910/DVN/LCYXPO}
\item \url{http://github.com/bzhan/bfh_python}
\item \url{http://regina-normal.github.io}
\end{itemize}


\begin{problems}
\item Get SageMath and SnapPy working together on your laptop, for
  example using the
  \href{http://bitbucket.org/t3m/sagedocker}{computop Docker
    image}.  Alternatively, from any of the physical ICERM terminals
  you can access it via \url{http://icerm2.icerm.brown.edu:8888}.

\item You can get a knot of 14 or fewer crossings in SnapPy by
  doing:

  \vspace{-0.5cm}
  
  \begin{verbatim}
  knots = snappy.HTLinkExteriors(cusps=1)
  E = knots.random()
  \end{verbatim}

  \vspace{-1cm}
  
  Use verified computation as described here:
  \url{http://snappy.computop.org/verify.html} \\
  to prove it is hyperbolic and to compute its volume to a provably
  correct 250 decimal places.  By Mostow rigidity this number is an
  invariant of the knot exterior and hence of the knot itself. (There
  are a handful of non-hyperbolic links in this range, so you're very
  unlike to pick one of them and so be unable to complete this problem!)

\item 
  Look at the documentation for HTLinkExteriors by typing
  \vspace{-0.5cm}
  
  \begin{verbatim}
  ?snappy.HTLinkExteriors
  \end{verbatim}

  \vspace{-1cm} to figure out how to pick a random \textbf{10
   crossing} knot.  Download the software of
  \\
  \url{http://doi.org/10.7910/DVN/LCYXPO}
  \\
  and use it to find coorientable taut/Reebless foliations on at least
  one Dehn surgery of your random knot.
  
\item Python programming practice: 

  Use \url{http://snappy.computop.org/spherogram.html} to write a
  Python function to produce a link projection of the
  $(a_1, a_2, \ldots, a_k)$ pretzel link.  For the $(-2, 3, 7)$
  pretzel knot, write a procedure that searches for the two slopes of
  the two len space Dehn surgeries discovered by Fintushel-Stern.  Use
  Regina to determine which lens spaces these are.  Can you find lens
  space surgeries on other pretzel knots?

\item Look at the list of software that is part of the computop Docker
  image.  See if you can compute something interesting with one of them.
  
\item The webpage \url{http://computop.org} lists a wide variety of
  computational tools in low\hyp dimensional topology.  Find one that
  is relevant to your own work and try to get it working in your
  Docker container.

\end{problems}
  
\end{document}